---
title: "General Social Survey"
author: "Ayinda St.Holder"
date: "April 27,2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"
 

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.

How are the "voted in 2016" and "voted in 2012" related?

$H_0:$ There's no relations between the two year 2016 and 2012.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Using the categorial variable, and a Chi- squared test to test the hypothesis that voting in 2016 is / is not independent of voting in 2012.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

Chi-squared statistic revealas a p- value of which 0 is less than 0.05. Therefore, we reject the null hypothesis. Bar Chart shows a clear relationship between voters in 2016 and 2012.

### Descriptive Results

We used a Bar Chart to show voting trends between 2016 and 2012.



#### Graphical Descriptive Results

```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent",xlab="voted in 2016") 

```

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Nummerical summary as a cross tabulation:
```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)
```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```
The hypothesis was voting in 2016 is related to 2012.
Chi-squared statistic revealas a p- value of which 0 is less than 0.05. Therefore, we reject the null hypothesis. Bar Chart shows a clear relationship between voters in 2016 and 2012. 

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

$H_0:$ Belief in the Big Bang is not related to belief in the afterlife.

$H_0:$ Belief in the Big Bang is related to belief in the after life.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Using catergorical variables and a Chi - squared test to test the hypothesis that belief in the Big Bang is / is not realeted to belief in the after life.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

We use the Bar Chart to show the relationship between belief in the Big Bang and belief in an afterlife.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent",xlab="Belief In The Big Bang") 

```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical results as a cross tabulation:
```{r}

xtabs(~BIGBANG+POSTLIFE, data=mydata3)

```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
bangtab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(bangtab)
```
The hypothesis believes in the Big Bang and believes in the Afterlife bare related.

Chi-squared statistic revealas a p- value of 0.0065  which is less than 0.05. Therefore, we must reject the null hypothesis and favor the alternative hypothesis that the belief in the bigbang IS related top belief in the afterlife.

 
# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

Are opinions on the death penalty and opinions in court related?

$H_0:$ Opinion on the penalty and opinion on courts ARE NOT related.  

$H_0:$ Opinion on the penalty and opinion on courts are realated. 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Use catergorical variables and a Chi squared test to test the hypothesis

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

We used a Bar Chart to show the opinion on the death penalty and the opinion on courts.

#### Graphical Descriptive Results
```{r}
barchartGC(~CAPPUN+COURTS, data=mydata4,type = "percent",xlab= 'view on CAPPUN',main='View on Courts')
```

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

Nummerical summary as cross tabulation:
```{r}
xtabs(~CAPPUN+COURTS, data=mydata4)
```

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata4)
captab=xtabs(~CAPPUN+COURTS, data=mydata4)
fisher.test(captab)
```
Hypothesis:
View on the death penalty and view on courts are related.

Chi- Squared statistic reveals a P-value of 0 which is less than 0.05. Therefore, we reject the null hypothesis in favor of the alternative hypothesis that the views on the death penalty and the views on courts are related.

# Discussion and Conclusion 

###Voting:
The hypothesis was voting in 2016 is related to 2012.

Chi-squared statistic revealas a p- value of which 0 is less than 0.05. Therefore, we reject the null hypothesis. Bar Chart shows a clear relationship between voters in 2016 and 2012. 

### Big Bang vs.Afterlife:
The hypothesis believes in the Big Bang and believes in the Afterlife bare related.

 Chi-squared statistic revealas a p- value of 0.0065  which is less than 0.05. Therefore, we must reject the null hypothesis and favor the alternative hypothesis that the belief in the bigbang IS related top belief in the afterlife.

###Cappun vs.Courts:

Hypothesis:

View on the death penalty and view on courts are related.

Chi- Squared statistic reveals a P-value of 0 which is less than 0.05. Therefore, we reject the null hypothesis in favor of the alternative hypothesis that the views on the death penalty and the views on courts are related.

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

